---
title: "Regression Model"
author: 'W241: Fox, George, McCleary'
date: "7/27/2022"
output: pdf_document
---

```{r load packages, message = FALSE, include = FALSE}
list.of.packages <- c("data.table", "sandwich", "lmtest", "ggplot2", "stargazer")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

library(data.table)
library(sandwich)
library(lmtest)
library(ggplot2) 
library(stargazer)
```

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
d <- data.table(readxl::read_xlsx("Final_Data_clean.xlsx"))
d[is.na(Value), account_disabled := T]
d[!is.na(Value), account_disabled := F]
d
```
```{r}
# Is there differential attrition between Facebook and Google?
website_diff_attrition_model <- d[, lm(account_disabled ~ as.factor(Website))]
website_diff_attrition_coeftest <- coeftest(website_diff_attrition_model, 
                                            vcovHC(website_diff_attrition_model))
```
```{r}
# Stargazer output.
stargazer(website_diff_attrition_model, header=F)
```

\begin{table}[!htbp] \centering 
  \caption{Differential Attrition between YouTube and Facebook} 
  \label{} 
\begin{tabular}{@{\extracolsep{5pt}}lc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{1}{c}{\textit{Dependent variable:}} \\ 
\cline{2-2} 
\\[-1.8ex] & Account Disabled \\ 
\hline \\[-1.8ex] 
  Website: YouTube & $-$0.413$^{***}$ \\ 
  & (0.031) \\ 
  & \\ 
 Constant & 0.413$^{***}$ \\ 
  & (0.022) \\ 
  & \\ 
\hline \\[-1.8ex] 
Observations & 504 \\ 
R$^{2}$ & 0.260 \\ 
Adjusted R$^{2}$ & 0.259 \\ 
Residual Std. Error & 0.349 (df = 502) \\ 
F Statistic & 176.378$^{***}$ (df = 1; 502) \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
\end{tabular} 
\end{table} 

```{r}
# Does demographic information influence Facebook's decision to ban an account?
facebook_diff_attrition_model <- d[Website=="Facebook", lm(account_disabled ~ as.factor(Gender) + as.factor(`Age Group`))]

facebook_diff_attrition_coeftest <- coeftest(facebook_diff_attrition_model, 
                                            vcovHC(facebook_diff_attrition_model))
```

```{r}
# Stargazer output.
stargazer(facebook_diff_attrition_model, header=F)
```

\begin{table}[!htbp] \centering 
  \caption{Differential Attrition on Facebook between Demographics} 
  \label{} 
\begin{tabular}{@{\extracolsep{5pt}}lc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{1}{c}{\textit{Dependent variable:}} \\ 
\cline{2-2} 
\\[-1.8ex] & Account Disabled \\ 
\hline \\[-1.8ex] 
 Gender: Male & $-$0.048 \\ 
  & (0.076) \\ 
  & \\ 
 Gender: Prefer Not To Answer & 0.071 \\ 
  & (0.076) \\ 
  & \\ 
 Age Group: Young & 0.063 \\ 
  & (0.062) \\ 
  & \\ 
 Constant & 0.373$^{***}$ \\ 
  & (0.062) \\ 
  & \\ 
\hline \\[-1.8ex] 
Observations & 252 \\ 
R$^{2}$ & 0.014 \\ 
Adjusted R$^{2}$ & 0.002 \\ 
Residual Std. Error & 0.493 (df = 248) \\ 
F Statistic & 1.177 (df = 3; 248) \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
\end{tabular} 
\end{table} 


Summary: Attrition is significantly different between YouTube and Facebook, however, Facebook attrition is not significantly different due to any of the collected demographic variables.
